\section{Introduction}
In recent years, forms of collaborations and cooperation in science have been transformed. As computing tools get generalized in larger scientific communities, the lab, which used to be studied in the classic Science Technology and Society (STS) literature \cite{latour1987science,houdart2008cour}, has been progressively replaced by new settings, redefining the standards \cite{galison1996disunity,kitchin2014big} on the way science is, and how it should be practiced \cite{calvert2013collaboration,leonelli2012introduction}. Additionally, while the last decades have been dominated by closed-source scientific software (e.g., MatLab, Mathematica), number of tools and libraries for research nowadays build upon community maintained open source programs and libraries, such as R, Numpy, Scipy, IPython Notebook, or Jupyter. These libraries are also increasingly publicly shared, for peer-review, for thorough reproduction of research results, and for reuse in other contexts and by other researchers. The lines between scientific software and libraries used for research are increasingly blurred, into the same complex ecosystem of code sharing. As a consequence of these emergent scientific practices, scientific tools are also increasingly achieved through the integration of heterogeneous perspectives and computational methods. In this context, the role of a researcher gets multiplied and diversified as he becomes a collaborator of his domain sciences, but also a contributor, a debugger, a user, and a promoter of a scientific library maintained by an open software community, in addition to more traditional tasks, such as conceiving research protocols, gathering data, testing hypotheses, and presenting results in scientific articles.

While open collaboration, as well as open and reproducible science, seems to be a genuine means and feature witnessed broadly in highly connected societies \cite{benkler2011penguin}, for the quantitative researcher, understanding the complex dynamics shaping self-organized communities has remained a challenge, despite the abundance of publicly available data available. Most often, large scale data mining quantitative studies lack sufficient context to extract the meaning of actions performed, while qualitative studies, focused on extracting context, might seem too partial and focused, thus preventing from getting a better big picture without losing context. 

However, the growing use of collaborative tools for scientific research has generated large amounts of data, which challenge the enduring lack of communication between qualitative and quantitative research methods.  As Venturini et al. recalled \cite{venturini2015fill}, traceability and contextualization of the data allow the emergence of a new form of quantitatively informed ethnography, particularly relevant to study the complex dynamics of scientific collaboration that social coding platforms, such as GitHub. Here, we address this methodological gap, by combining state-of-the-art modeling of social dynamics, with ethnography to analyze how a small group of scientists promote more open and reproducible science practices, in order to build epistemic communities and collaborative platforms in academia \cite{calvert2013collaboration}.

In this paper, we focus on a specific event, the Astro Hack Week, which was held at the University of Washington, from September 15-19, 2014. A large quantity of events that punctuated the Astro Hack Week, have been recorded on GitHub \cite{github}, a social coding platform, which rapidly transforms the way in which scientists are collaborating \cite{gerson2013integration}. Social coding platforms are promoting new ways to work together in data driven environments, where science is being redefined with some value of openness and reproducibility \cite{ducheneaut2005socialization}, including code accessibility and verification \cite{hayden2015rule,proebsting2015repeatability}. Looking at the dynamics of GitHub contributions by Astro Hack Week actors, and framing them in the socio-technical context in which they have emerged, we show that past the initial impulse, the Astro Hack Week triggered mid- term and long-term follow-up activity. We illustrate how this activity is only qualitatively tangentially related to the Astro Hack Week, and rather stems from the exposure to a new practice of science, and from the build-up of new social ties.

The rest of this paper is organized as follows. We first further expose the reader to the literature in relation with the emergence of platforms and tools, tuned for a rather self-organized practice of science, coming from the complex dynamics of computer supported collaborative work (CSCW) and Science and Technologies Studies (STS). We then present a method, which combines qualitative and quantitative analysis and in-depth modeling of actions taken by participants, before, during and after the Astro Hack Week. Data employed and results are then presented, and discussed. We finally conclude with limitations and future research directions.

%The way open and reproducible science evolves is a challenge, if not a paradigm change, for STS studies. In recent years, however, the development of collaborative tools for scientific research has generated a large amount of data that may help bridging the gap between qualitative and quantitative methods. As Tommaso et al. \cite{tommaso2015} {\bf [missing reference for Tommaso]} recalled, both the phenomenon of traceability and the capacity to re contextualized the data that have been collected allow the emergence of a new form of ethnography, which is both qualitatively and quantitatively informed. Such method seems particularly relevant to study the complex dynamics of scientific collaboration that social coding platforms, such as GitHub, are making possible and visible in the context of computational intensive scientific projects. 


%Like for most instances of open collaboration, academia needs to build a strong and large community of scientists. Important efforts have been undertaken across fields by universities (e.g., Software Carpentry \cite{software_carpentry}, or by discipline (across institutions) \cite{calvert2013collaboration}. In this paper, we focus on the Astro Hack Week (AstroWeek), which was held at the University of Washington, from September 15-19, 2014, with the support the Moore Foundation, the Sloan Foundation, and the UW eScience Institute. A large quantity of events that punctuated the AstroWeek, have been recorded on GitHub \cite{github}, a social coding platform, which rapidly transforms the way in which scientists are collaborating \cite{gerson2013integration}. GitHub is promoting new way to collaborate in data driven scientific environment where science is being redefined with some value of openness and reproducibility \cite{ducheneaut2005socialization} -- which include code accessibility  and verification \cite{hayden2015rule,proebsting2015repeatability}.

%Looking at the dynamics of contributions by actors, we show that past the initial impulse, AstroWeek triggered long-term follow-up activity. We illustrate how this long-term activity is only qualitatively tangentially related to the AstroWeek, and rather stems from the exposure to a new practice of science, and from the build-up of new social ties. We also document the ``costs"  of senior data scientists spending time teaching data science to lay participants.



%During the days the time was divided between the sessions dedicated of teaching and learning of coding, statistics, and skills required for data analysis of large data set and the afternoon where dedicated to specific projects and breakout sessions. 


%To address the problem of community building and onboarding, we study the community of astronomers {\bf [astrophysicists? cosmologists?]}, learning and practicing data science. In particular, we focus on the Astro Hack Week, which was held at the University of Washington on September 15-19, 2014 (with the support the Moore Foundation, the Sloan Foundation, and the UW eScience Institute). 


%Here, we shall document how a one-week hackathon (the Astro Hack Week), organized for the purpose of engaging scientists in the practice of open and reproducible data science, has short-, mid-term and long-term spillovers {\bf [not sure for the long-term]}, as a result of critical cascades of repository creations and forks \cite{}, as well as contributions to broader data science community {\bf [it's not clear here, but the goal is to convey the idea that while projects during the Astro Hack Week, may be short lived, the mere fact of having people interact (physically), creates opportunities for future stand-alone contributions or even collaborations (c.f., testimony of Kyle)]}

